---
title: "PISCO_kelp_urchin_swath.Rmd"
author: "Andrés Pinos-Sánchez"
date: "2025-07-24"
output: 
  html_document: default
  pdf_document: default
---

#### AIM OF CODE

Obtain for bull kelp and urchin metrics for ORTM model parameters. Here we use
PISCO swath and tuffy datasets filtered for the Oregon coast and target species
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 

# Load libraries
library(ggplot2)
library(dplyr)
library(readr)
library(tidyverse)
library(ggpubr)
library(brms)
library(posterior)
library(coda)

# Clear workspace and set seed
rm(list = ls()) #environment
set.seed(123)

# Working directory
# setwd("C:/Users/pinosa/Documents/Git/OR_trophic_model/Code/Data_DataExploration/PISCO_kelp_urchin_swath_data")

```

#### GETTING THE DATA

This data is from PISCO surveys
O.G. data: https://data.piscoweb.org/metacatui/view/doi%3A10.6085%2FAA%2FPISCO_kelpforest.1.10
```{r}
# Read PISCO SWATH data
taxon <- read_csv("PISCO_kelpforest_taxon_table.1.3.csv") #taxon table
sites <- read_csv("PISCO_kelpforest_site_table.1.3.csv") #sites table
swath <- read_csv("PISCO_kelpforest_swath.1.3.csv") #Swath data
quad  <- read_csv("PISCO_kelpforest_quad.1.3.csv") #Quad data
fish  <- read_csv("PISCO_kelpforest_fish.1.4.csv") #fish data

# Read PISCO TUFFY data
tuffy <- read_csv("PISCO_tuffy_all_years.csv") #Tuffy data (compiled by Andres)

```

Annotate MPA status (only relevant for CA model)
```{r}
# Step 1: Create site_status variable
site_status <- sites %>% select(site, site_status) %>% distinct()

# Step 2: Insert site_status in swath and quad data sets (relevant for MPAs)
swath <- left_join(swath, site_status, by = "site")
quad  <- left_join(quad,  site_status, by = "site")

```

Create an Oregon and region (north & south) filter to be used in the swath data
```{r}
# Step 1: Filter for Oregon sites (>=42.0 lat) and assign regions (north & south)
oregon_region <- sites %>%
  filter(latitude >= 42.0) %>%
  mutate(region = case_when(latitude >= 42.0 & latitude < 43.5 ~ "South",
                            latitude >= 44.25                 ~ "North"))

# Step 2: Apply filter and join region info
oregon_region <- oregon_region %>%
  select(site, region) %>%
  distinct(site, .keep_all = TRUE)

```


#### BULL KELP ADULTS

Determine: Scaling transition (spatial variance in kelp densities)
```{r}
# Step 1: Filter for Bull kelp (NERLUE) from swath data
kelp_adult <- swath %>%
  filter(classcode == "NERLUE") %>%
  select(year, month, site, zone, transect, count, size, depth)

# Step 2: Filter kelp_adult data to Oregon sites and join region info
kelp_adult <- kelp_adult %>%
  filter(site %in% oregon_region$site) %>%
  left_join(oregon_region, by = "site")

# Step 3: Fill missing depth values

  ## Fill using site + zone mean
  kelp_adult <- kelp_adult %>%
    group_by(site, zone) %>%
    mutate(depth = if_else(is.na(depth),
                           mean(depth, na.rm = TRUE),
                           depth)) %>%
    ungroup()
  
  ## Fill remaining with global mean
  kelp_adult <- kelp_adult %>%
    mutate(depth = if_else(is.na(depth),
                           mean(depth, na.rm = TRUE),
                           depth))

# Step 4: Calculate total biomass per entry
kelp_adult <- kelp_adult %>%
  mutate(size = depth,
         biomass = 0.0346 * exp(0.501 * size) * count) #(Stekoll et al., 2007)

# Step 5: Scaling transitions (overall and by region)

  # OVERALL scaling transition (Oregon coast)
  scaling_transition_OR <- kelp_adult %>%
    group_by(year, site, zone) %>%
    summarise(biomass_var = var(biomass, na.rm = TRUE), .groups = "drop") %>%
    group_by(year) %>%
    summarise(mean_var = mean(biomass_var, na.rm = TRUE), .groups = "drop") %>%
    summarise(scaling_transition = mean(mean_var, na.rm = TRUE))

  # BY REGION scaling transitions
  scaling_transition_by_region <- kelp_adult %>%
    group_by(region, year, site, zone) %>%
    summarise(biomass_var = var(biomass, na.rm = TRUE), .groups = "drop") %>%
    group_by(region, year) %>%
    summarise(mean_var = mean(biomass_var, na.rm = TRUE), .groups = "drop") %>%
    group_by(region) %>%
    summarise(scaling_transition = mean(mean_var, na.rm = TRUE), .groups = "drop")
  
# Results:
scaling_transition_OR  # overall transition
scaling_transition_by_region  # regional transitions

```


#### BULL KELP RECRUITS

Determine: Temporal (norm) variation|standard deviation (noise) of kelp recruits. Note: we are assuming kelp on month "8" (August) are all successfully settle recruits 
```{r}
# Step 1: Filter for Bull kelp (NERLUE) from swath data for August (8)
kelp_recruit <- swath %>%
  filter(classcode == "NERLUE", month == 8) %>%
  select(year, month, site, zone, transect, count)

# Step 2: Filter kelp_recruit to Oregon sites and join region info
kelp_recruit <- kelp_recruit %>%
  filter(site %in% oregon_region$site) %>%
  left_join(oregon_region, by = "site")

# Step 3: Calculate mean successful recruits across space and time
kelp_recruit <- kelp_recruit %>%
  group_by(year, site, region) %>% 
  summarise(mean_count = mean(count), .groups = "drop")

# Step 4: GLM to detrend site and year effects & Normalize and calculate sd (temporal noise)
model_rec <- glm(mean_count ~ year * site, data = kelp_recruit)
ts_detrended <- mean(kelp_recruit$mean_count) + residuals(model_rec)
noise_rec     <- sd(ts_detrended / mean(ts_detrended))

# Step 5: GLM per region to detrend site and year effects & Normalize and calculate sd (temporal noise)
noise_rec_by_region <- kelp_recruit %>%
  group_by(region) %>%
  summarise(noise = {
    model <- glm(mean_count ~ year * site, data = cur_data())
    ts_detrended <- mean(cur_data()$mean_count) + residuals(model)
    sd(ts_detrended / mean(ts_detrended))},
    .groups = "drop")

# Results:
noise_rec # overall variation (noise)
noise_rec_by_region # regional variation (noise)

```


#### URCHINS ADULTS

Determine: Spatio-temporal variance on adult urchin biomass
```{r}
# Step 1: Biomass constants (Ling et al. 2015)
red_bio <- 0.0499 * 7   + 0.0019
pur_bio <- 0.0499 * 4.1 + 0.0019

# Step 2: Aggregate counts per transect & compute biomass
urchin_adult <- swath %>%
  filter(classcode %in% c("MESFRAAD", "STRPURAD")) %>%
  select(year, site, zone, transect, classcode, count, size) %>%
  filter(site %in% oregon_region$site) %>%
  left_join(oregon_region, by = "site") %>%
  group_by(year, site, zone, transect, classcode) %>%
  summarise(count = sum(count), .groups = "drop") %>%
  pivot_wider(names_from = classcode,
              values_from = count,
              values_fill = 0) %>%
  transmute(year, site, zone, transect,
            red_count    = MESFRAAD,
            purp_count   = STRPURAD,
            red_biomass  = red_count  * red_bio,
            purp_biomass = purp_count * pur_bio,
            total_biomass = red_biomass + purp_biomass)

# Step 3: Spatial variance in biomass
urchin_var <- urchin_adult %>%
  pivot_longer(cols = c(red_biomass, purp_biomass, total_biomass),
               names_to = "spp",
               values_to = "biomass") %>%
  group_by(year, site, zone, spp) %>%
  summarise(var = var(biomass, na.rm = TRUE), .groups = "drop")

# Step 4: Temporal and overall metrics
urchin_yearvar <- urchin_var %>%
  group_by(year, spp) %>%
  summarise(mean_var = mean(var, na.rm = TRUE), .groups = "drop") %>%
  group_by(spp) %>%
  summarise(year_var = mean(mean_var, na.rm = TRUE))

urchin_overall <- urchin_var %>%
  group_by(spp) %>%
  summarise(overall_var = mean(var, na.rm = TRUE))

# Results:
urchin_yearvar
urchin_overall # target number

```


#### URCHIN RECRUITS

Determine: Temporal (norm) variation|standard deviation (noise) of sea urchin recruits.Note: SWATH data doesn't contain urchin recruits info for OR, therefore we are going to use the PISCO TUFFY data set
```{r}
# Step 1: Filter for urchin recruits taxa and remove unwanted sites (keep OR only)
urchin_recruit <- tuffy %>%
  filter(count_classcode %in% c("66", "67", "70"),
         !site_code %in% c("CMEN00", "CMES00", "IFBRXX", "KHLX00",
                           "PSGX00", "SHCX00", "TRHX00", "IPTAXX"))

# Step 2: Group to site-year recruit totals
urchin_recruit_sy <- urchin_recruit %>% #site-year "sy"
  group_by(site_code, year) %>%
  summarise(count = sum(count), .groups = "drop")

# Step 3: Fit a zero‑inflated Poisson with site and nested year random effects
brms_rec <- brm(formula = count ~ 1 + (1 | site_code / year), 
                #count ~ site_code + (1 | site_code / year)
                family  = zero_inflated_poisson(),
                data    = urchin_recruit_sy,
                iter    = 4000,
                warmup  = 1000,
                thin    = 3,
                chains  = 3,
                cores   = 3)

summary(brms_rec)
plot(brms_rec)

# Result: Extract posterior draws for the year-within-site sd parameter
mean(as_draws_df(brms_rec)$`sd_site_code:year__Intercept`)

```


#### KELP AND URCHINS

Plot urchin vs kelp biomass
```{r}
# Step 1: Join kelp and urchin tables
KU_OR <- kelp_adult %>% 
  full_join(urchin_adult, by = c("year", "site", "zone", "transect"))

# Step 2: Compare kelp vs urchin biomasses

  ## Plot purple urchins only
  ggplot(KU_OR) +
    geom_point(aes(x = purp_biomass, y = biomass, colour = as.factor(year))) +
    ylim(0, 2000) +
    theme_bw() +
    labs(x = "Purple urchin biomass (kg·60m²)", 
         y = "Bull kelp biomass (kg·60m²)", 
         colour = "Year")
  
  ## Plot red urchins only
  ggplot(KU_OR) +
    geom_point(aes(x = red_biomass, y = biomass, colour = as.factor(year))) +
    ylim(0, 2000) +
    theme_bw() +
    labs(x = "Red urchin biomass (kg·60m²)", 
         y = "Bull kelp biomass (kg·60m²)", 
         colour = "Year")
  
  ## Reshape table to compare both urchins
  KU_OR1 <- KU_OR %>% 
    select(year:transect, biomass, red_biomass, purp_biomass) %>%
    pivot_longer(cols = c(red_biomass, purp_biomass), 
                 names_to = "urchin_type", 
                 values_to = "urchin_biomass") %>% 
    mutate(urchin_type = str_extract(urchin_type, ".*(?=(_))")) %>% 
    mutate(urchin_type = factor(urchin_type, 
                                levels = c("purp", "red"), 
                                labels = c("Purple", "Red")))
  
  ## Plot both urchin types
  ggplot(KU_OR1) +
    geom_point(aes(x = urchin_biomass, y = biomass, colour = urchin_type)) +
    theme_bw() +
    labs(x = "Urchin biomass (kg·60m²)", 
         y = "Bull kelp biomass (kg·60m²)", 
         colour = "Urchin type") +
    ylim(0, 500)

```
# Determine: kelp biomass density threshold in which a barren state is declared. Choose most healthy kelp forest and the one withn longest sampling in time. We are choosing REDFISH because of its MPA status, and because it is the closest to Coos Bay which is the healthiest kelp forest in Oregon according to the ORKA report
```{r}
# Summary statistics for OTTER and REDFISH sites
KU_OR %>%
  filter(str_starts(site, "OTTER") | str_starts(site, "REDFISH")) %>%
  mutate(site_group = case_when(
    str_starts(site, "OTTER") ~ "OTTER",
    str_starts(site, "REDFISH") ~ "REDFISH"
  )) %>%
  group_by(site_group) %>%
  summarise(across(c(biomass, total_biomass),
                   list(mean = ~mean(.x, na.rm = TRUE),
                        sd = ~sd(.x, na.rm = TRUE)))) -> summary_stats

print(summary_stats)

# Time series plot of mean kelp biomass by site group
KU_OR %>%
  filter(str_starts(site, "OTTER") | str_starts(site, "REDFISH")) %>%
  mutate(site_group = case_when(
    str_starts(site, "OTTER") ~ "OTTER",
    str_starts(site, "REDFISH") ~ "REDFISH"
  )) %>%
  group_by(year, site_group) %>%
  summarise(mean_biomass = mean(biomass, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_biomass, color = site_group)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  labs(x = "Year",
       y = "Mean Bull Kelp Biomass (kg/60m²)",
       title = "Mean Bull Kelp Biomass Over Time at OTTER and REDFISH Sites",
       color = "Site Group")

```
